The emotion--facial expression link: Evidence from human and automatic expression recognition

### 2019-11-06 manuscript_apa.Rmd
manuscript doc

# output: papaja::apa6_word

output            :
  bookdown::pdf_book:
    toc: false
    number_sections: false
    base_format: papaja::apa6_pdf

@@

This annotation procedure resulted in a uni-dimensional time-series for each video per human observer, identifying for each second of the video which emotion was recognized. Then, time-series corresponding to the same video were aggregated to calculate the proportion of human observers $x_{video_{i}.label_{j}.t_{k}}$ for each second of the video per emotional label (EQ\ref{eq:1}).

\begin{equation}
\label{eq:1}
x_{video_{i}.label_{j}.t_{k}} = \frac{n_{video_{i}.label_{j}.t_{k}}}{n_{video_{k}}}
\end{equation}

@@

For both human and automatic recognition, to determine which of the six basic emotions can be used to identify each video, the recognition probability for each label by frame was converted into odd ratio by label [@dente2017measures]. The highest sum of each odd ratio time-series defines the label recognized by the automatic classifier (EQ\ref{eq:2} for human recognition and EQ\ref{eq:3} for automatic recognition).

\begin{equation}
\label{eq:2}
video_{i}.label = \max\left(\frac{\sum_{k=1}^{n}x_{video_{i}.label_{j}.t_{k}}}{\sum_{k=1}^{n}x_{video_{i}.t_{k}}}\right)
\end{equation}

\begin{equation}
\label{eq:3}
video_{i}.label = \max\left(\frac{\sum_{k=1}^{f}p_{video_{i}.label_{j}.t_{k}}}{\sum_{k=1}^{f}p_{video_{i}.t_{k}}}\right)
\end{equation}

@@

recognition_result <- grid_data %>%
  dplyr::left_join(self_report_result, by = c(
    "list_video" = "C_Video",
    "list_emotion" = "sr_emotion"
    )
  ) %>%
  dplyr::left_join(human_recognition_result, by = c(
    "list_video" = "C_Video",
    "list_emotion" = "hr_emotion"
    )
  ) %>%
  dplyr::left_join(automatic_recognition_result, by = c(
    "list_video" = "C_Video",
    "list_emotion" = "ar_emotion"
    )
  ) %>%
  replace(is.na(.), 0)

@@

```{r confusionTable-sr-hr, results="asis"}
confusionMatrix_sr_hr_table <- confusionMatrix_sr_hr %>%
  magrittr::use_series(byClass) %>%
  as.data.frame() %>%
  tibble::rownames_to_column("Emotion") %>%
  dplyr::mutate(Emotion = gsub(pattern = "Class: ",replacement = "",x = Emotion)) %>% 
  dplyr::select(Emotion, Sensitivity, Specificity, Precision, F1) %>% 
  dplyr::mutate_if(is.numeric,round,2) %>% 
  replace(is.na(.), "\\textit{na.}")

confusionMatrix_sr_hr_table %>%
  knitr::kable(
    "latex",
    caption = "(ref:confusionTable-sr-hr)",
    booktabs = TRUE,
    digits = 2,
    linesep = "",
    escape = FALSE,
    align = c('l', rep('c', 4)),
    longtable = T
  ) %>%
  kableExtra::kable_styling(
    latex_options = "HOLD_position",
    font_size = 8,
    full_width = TRUE 
  ) %>%
  kableExtra::footnote(
    general_title = "",
    general = "Note. na. values are produced when not enough data are available to compute accuracy indicators.", 
    escape = FALSE,
    footnote_as_chunk = TRUE,
    threeparttable = TRUE
  )
```

@@

```{r confusionTable-sr-ar, results="asis"}
confusionMatrix_sr_ar_table <- confusionMatrix_sr_ar %>%
  magrittr::use_series(byClass) %>%
  as.data.frame() %>%
  tibble::rownames_to_column("Emotion") %>%
  dplyr::mutate(Emotion = gsub(pattern = "Class: ",replacement = "",x = Emotion)) %>% 
  dplyr::select(Emotion, Sensitivity, Specificity, Precision, F1) %>% 
  dplyr::mutate_if(is.numeric,round,2) %>% 
  replace(is.na(.), "\\textit{na.}")

confusionMatrix_sr_ar_table %>%
  knitr::kable(
    "latex",
    caption = "(ref:confusionTable-sr-ar)",
    booktabs = T,
    digits = 2,
    linesep = "",
    escape = FALSE,
    align = c('l',rep('c', 4)),
    longtable = T
  ) %>%
  kableExtra::kable_styling(
    latex_options = c("HOLD_position"),
    font_size = 8,
    full_width = TRUE
    ) %>%
  kableExtra::footnote(
    general_title = "",
    general = "Note. na. values are produced when not enough data are available to compute accuracy indicators.", 
    escape = FALSE,
    footnote_as_chunk = TRUE,
    threeparttable = TRUE
    )
```

@@

confusionMatrix_hr_ar_freq %>% 
  ggplot(mapping = aes(x = hr_emotion, y = ar_emotion)) +
  geom_tile(aes(fill = prop_hr), colour = "white") +
  geom_text(
    aes(label = paste0(scales::percent(prop_hr), "\n(", Freq,"/",n_hr, ")")),
    color = "black",
    size = 4,
    family="serif",
    parse = FALSE,
    lineheight = 0.7)+
  scale_fill_gradient(
    name = "Proportion Recognized Emotion",
    low = "white", 
    high = "red", 
    labels = scales::percent,
    limits = c(0,1)
  ) +
  scale_x_discrete(name = "Human Observers") +
  scale_y_discrete(name = "Automatic Classifier") +
  theme_minimal() +
  theme(text = element_text(size=16,family="serif"),
        axis.text.x = element_text(size=16, angle = 45, hjust = 0.75,vjust=0.9),
        axis.text.y = element_text(size=16, hjust = 0.5),
        axis.title.y = element_text(margin = margin(t = 0, r = 20, b = 0, l = 0)),
        legend.text = element_text(size=8),
        legend.position = "bottom")
```

>>>>

output: papaja::apa6_word

# output            :
#   bookdown::pdf_book:
#     toc: false
#     number_sections: false
#     base_format: papaja::apa6_pdf

@@

This annotation procedure resulted in a uni-dimensional time-series for each video per human observer, identifying for each second of the video which emotion was recognized. Then, time-series corresponding to the same video were aggregated to calculate the proportion of human observers $x_{video_{i}.label_{j}.t_{k}}$ for each second of the video per emotional label shown in Equation \@ref(eq:one).

\begin{equation}
x_{video_{i}.label_{j}.t_{k}} = \frac{n_{video_{i}.label_{j}.t_{k}}}{n_{video_{k}}}(\#eq:one)
\end{equation}

@@

For both human and automatic recognition, to determine which of the six basic emotions can be used to identify each video, the recognition probability for each label by frame was converted into odd ratio by label [@dente2017measures]. The highest sum of each odd ratio time-series defines the label recognized (see Equation \@ref(eq:two) for human recognition and Equation \@ref(eq:three) for automatic recognition).

\begin{equation}
video_{i}.label = \max\left(\frac{\sum_{k=1}^{n}x_{video_{i}.label_{j}.t_{k}}}{\sum_{k=1}^{n}x_{video_{i}.t_{k}}}\right)(\#eq:two)
\end{equation}

\begin{equation}
video_{i}.label = \max\left(\frac{\sum_{k=1}^{f}p_{video_{i}.label_{j}.t_{k}}}{\sum_{k=1}^{f}p_{video_{i}.t_{k}}}\right)(\#eq:three)
\end{equation}

@@

recognition_result <- grid_data %>%
  dplyr::left_join(
    self_report_result,
    by = c(
      "list_video" = "C_Video",
      "list_emotion" = "sr_emotion"
      )
    ) %>%
  dplyr::left_join(
    human_recognition_result,
    by = c(
      "list_video" = "C_Video",
      "list_emotion" = "hr_emotion"
      )
    ) %>%
  dplyr::left_join(
    automatic_recognition_result,
    by = c(
      "list_video" = "C_Video",
      "list_emotion" = "ar_emotion"
      )
    ) %>%
  replace(is.na(.), 0)

@@

```{r confusionTable-sr-hr, results="asis"}
confusionMatrix_sr_hr_table <- confusionMatrix_sr_hr %>%
  magrittr::use_series(byClass) %>%
  as.data.frame() %>%
  tibble::rownames_to_column("Emotion") %>%
  dplyr::mutate(Emotion = gsub(pattern = "Class: ",replacement = "",x = Emotion)) %>% 
  dplyr::select(Emotion, Sensitivity, Specificity, Precision, F1) %>% 
  dplyr::mutate_if(is.numeric,round,2) %>% 
  # replace(is.na(.), "\\textit{na.}")
  replace(is.na(.), "na.")

confusionMatrix_sr_hr_table %>%
  # knitr::kable(
  #   "latex",
  #   caption = "(ref:confusionTable-sr-hr)",
  #   booktabs = TRUE,
  #   digits = 2,
  #   linesep = "",
  #   escape = FALSE,
  #   align = c('l', rep('c', 4)),
  #   longtable = T
  # ) %>%
  # kableExtra::kable_styling(
  #   latex_options = "HOLD_position",
  #   font_size = 8,
  #   full_width = TRUE 
  # ) %>%
  # kableExtra::footnote(
  #   general_title = "",
  #   general = "Note. na. values are produced when not enough data are available to compute accuracy indicators.", 
  #   escape = FALSE,
  #   footnote_as_chunk = TRUE,
  #   threeparttable = TRUE
  # )
  papaja::apa_table(
    caption = "(ref:confusionTable-sr-hr)",
    note = "na. values are produced when not enough data are available to compute accuracy indicators.",
    escape = TRUE
  )
```

@@

```{r confusionTable-sr-ar, results="asis"}
confusionMatrix_sr_ar_table <- confusionMatrix_sr_ar %>%
  magrittr::use_series(byClass) %>%
  as.data.frame() %>%
  tibble::rownames_to_column("Emotion") %>%
  dplyr::mutate(Emotion = gsub(pattern = "Class: ", replacement = "", x = Emotion)) %>% 
  dplyr::select(Emotion, Sensitivity, Specificity, Precision, F1) %>% 
  dplyr::mutate_if(is.numeric, round, 2) %>% 
  #replace(is.na(.), "\\textit{na.}")
  replace(is.na(.), "na.")

confusionMatrix_sr_ar_table %>%
  # knitr::kable(
  #   "latex",
  #   caption = "(ref:confusionTable-sr-ar)",
  #   booktabs = T,
  #   digits = 2,
  #   linesep = "",
  #   escape = FALSE,
  #   align = c('l',rep('c', 4)),
  #   longtable = T
  # ) %>%
  # kableExtra::kable_styling(
  #   latex_options = c("HOLD_position"),
  #   font_size = 8,
  #   full_width = TRUE
  #   ) %>%
  # kableExtra::footnote(
  #   general_title = "",
  #   general = "Note. na. values are produced when not enough data are available to compute accuracy indicators.", 
  #   escape = FALSE,
  #   footnote_as_chunk = TRUE,
  #   threeparttable = TRUE
  #   ) 
  papaja::apa_table(
    caption = "(ref:confusionTable-sr-ar)",
    note = "na. values are produced when not enough data are available to compute accuracy indicators.",
    escape = TRUE
  )
```

@@

confusionMatrix_hr_ar_freq %>% 
  ggplot(mapping = aes(x = hr_emotion, y = ar_emotion)) +
  geom_tile(aes(fill = prop_hr), colour = "white") +
  geom_text(
    aes(label = paste0(scales::percent(prop_hr), "\n(", Freq, "/", n_hr, ")")),
    color = "black",
    size = 4,
    family="serif",
    parse = FALSE,
    lineheight = 0.7)+
  scale_fill_gradient(
    name = "Proportion Recognized Emotion",
    low = "white", 
    high = "red", 
    labels = scales::percent,
    limits = c(0,1)
  ) +
  scale_x_discrete(name = "Human Observers") +
  scale_y_discrete(name = "Automatic Classifier") +
  theme_minimal() +
  theme(text = element_text(size=16,family="serif"),
        axis.text.x = element_text(size=16, angle = 45, hjust = 0.75,vjust=0.9),
        axis.text.y = element_text(size=16, hjust = 0.5),
        axis.title.y = element_text(margin = margin(t = 0, r = 20, b = 0, l = 0)),
        legend.text = element_text(size=8),
        legend.position = "bottom")
```
